### 2021-04-07 paper/filters.tex

Filter synthesis translates the contiguity type recognizer into CakeML in a way that preserves the precision of the recognizer to guarantee that the CakeML recognizer recognizes the same language as the contiguity type recognizer from which it is derived. This statement is the main result of the paper for filters.