### 2022-03-14 paper/semantics.tex

Brief, high-level overview, of the AGREE semantics for component synthesis.
These semantics are those defined in HOL.
The component is defined by its state.
The state is its internal state and current output.
Input streams are constrained to obey the assumptions.
At each input, the \textbf{eq}-statements are evaluated, in order, to update the internal state and outputs.
After the \textbf{eq}-statements are evaluated, guarantees are checked.
The section should define clearly what it means for a component to be \emph{well-formed}. 
The section should conclude by stating the \emph{Component Correct} definition.

The correctness, and structure, of the synthesized CakeML code is based on these semantics.

\subsection{Old AGREE Semantics}
\label{agree-semantics}

Here the formal semantics of the AGREE contract specification are
briefly presented to make clear the meaning of a high-assurance
component.  These semantics are the basis of proofs showing that a
synthesized component has input to output behavior specified by its
contract.

Assume that all data flowing between components is in the form of byte
arrays, thus input and output is modelled at the level of what is
passed ``over the wire'' by the communication fabric. An \emph{environment},
$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to
strings where an L-value is anything that can appear on the left-hand
side of an assignment such as a named port, a field in a record, a
entry in an array, a local value defined by an eq-statement etc.

Let $s$ be an AGREE contract specification for some high-assurance
component. The notation $\theta_s$ is used to denote the environment that
contains a binding for any L-value in the scope of $s$ and nothing else, and
the notation $\Theta_s$ denotes the universe of all such environments. The
semantics of $s$ are defined over streams, $\pi = \theta_1, \theta_2, \ldots$, which
are finite sequences of environments, $\pi \in \Theta_s^*$.

%% \begin{comment}
%% [\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
%% generate a contig type for the inputs plus state vars.} ]

%% [Question: what is a \emph{total} environment? One that supplies at
%% least the bindings in $\theta$ needed to evaluate the spec?]
%% \end{comment}

The function $\konst{eval}\; s\; \pi$ evaluates $s$ on the stream,
$\pi$, and returns \konst{true} if $s$ is invariant along the entire
stream and \konst{false} otherwise. A guarantee $\mathcal{G}$ in $s$
is \emph{invariant} if $\mathcal{G}$ is true for each prefix of $\pi$,
while an eq-statement in $s$ is invariant if its binding in the
context of every step is equivalent to the computed value of its
associated expression in that same context. All guarantees and
eq-statements must be invariant in the stream for the function to
return \konst{true}.

The meaning of an AGREE contract specification is now defined as the
set of environment streams on which it is invariant.
%
\[
  \Lang{s} = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
\]
%
\noindent Intuitively, any stream $\pi \in \mathcal{L}(s)$, at each step, binds
the L-values in the eq-statements in a way that is consistent with
their associated expressions and the guarantees are all true in that
same step.

We claim that a synthesized high-assurance component preserves the
input to output relationship in the specification $s$ over every
stream in $\Lang{s}$.  Let $\pi^\prime = \konst{SynthEval}\;
s\; \pi$ denote a function that synthesizes $s$ and then evaluates that
synthesized component on the stream $\pi$ to create a new stream
$\pi^\prime$ containing added output and other bindings. We say that
two streams are equivalent in regards to a specification, denoted as
$\pi =_s \pi^\prime$, if and only if the two streams are the same
length and agree on bindings for the input and output for $s$ at every
step of the streams.  We now formally state the correctness claim for
synthesis.
%
\[
  \forall \pi \in \Lang{s}.\; (\konst{SynthEval}\; s\; I_s(\pi)) =_s \pi\\
\]
%
where $I_s(\pi)$ returns the corresponding stream that only retains
bindings for inputs in each step and nothing else. The claim is that
the synthesized component generates the same output stream as the
specification for any stream belonging to the specification that is
restricted to just input bindings at each step.