### 2022-03-14 paper/agree.tex
Revise example and add TODOs in all sections

The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components.
The AGREE specifications generated by the transforms for the high-assurance components are also explained.

\newsavebox{\sw}
\begin{lrbox}{\sw}
\begin{lstlisting}[style=agree,numbers=left]
eq req : bool = event(AutomationRequest);
eq avl : bool = event(AirVehicleLocation);
eq wp : bool = event(Waypoint);
eq strt: bool = event(Start);
eq alrt : bool = event(Alert);

assume "Automation requests are well-formed" :
  req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
assume "Air vehicle locations are well-formed" :
  avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);    
assume "One automation request in flight at a time" :
  true -> 
  (req => pre(Historically(not req) or Since(not req, strt)));

guarantee "Waypoints coincide with air vehicle locations":
  wp => avl;
guarantee "Starts include a new waypoint" :
  strt => wp;
guarantee "Waypoints are well-formed" : 
  wp => WELL_FORMED_WAYPOINT(Waypoint);
guarantee "Starts within one cycle of requests if not alerting" :
  (strt => ((not alrt) and req)) -> 
  (strt => ((not alrt) and (req or pre(req))));
guarantee "Alert if not started within one cycle of requests" :
    true -> ((pre(req and not strt) and not strt) => alrt);
guarantee "Once alerted always alerted" :
  not alrt or (Once(alrt) and alrt);
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \scalebox{0.62}{\usebox{\sw}}
  \end{center}
  \caption{The SW component contract.}
  \label{fig:sw}
\end{figure}

The AGREE specification for the SW component in the example of Section~\ref{sec:example} with the added cyber requirements is given in \figref{fig:sw}.
The AGREE specification language is a first-order predicate calculus that uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}.
As with Lustre, the semantics are synchronous dataflow where the inputs, outputs, and expressions are data streams that comply with the input assumptions.
Contracts are evaluated in dependency order with inputs being propagated to outputs through all contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic.\footnote{An apparent syntactic cycle, where a component is linked back to itself, may be broken temporally by inserting delay elements.}
Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream.
The semantics do not model computation or communication delay.
The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

The AGREE model checker attempts to prove several properties of the top-level model being verified.
The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component as well as to satisfy the guarantees of the output of the top-level component being verified (\ie, the system composition meets input assumptions at each input as well as the guarantees on the system output).
These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).
The next set of properties prove that the contract specifications for each component are \emph{self-consistent} \ie, a contract does not contradict itself.
These are the unexpanded results at the bottom of the figures.

Returning back to the contract of \figref{fig:sw}, it uses \texttt{eq} statements to define variables local to the contract specification.
For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression.
In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component.
The semantics used here do not buffer these events so the implicit input (or output) is only a boolean value.
An \texttt{event} expression refers to that implicit input (or output) and is true when data is placed on the named port.
The system contract here states assumptions on well-formed input, followed by guarantees on properties about the output.

REPLACE ALL: 
The \emph{Alert if start is not bounded relative to a request} guarantee is an invariant on the expression \texttt{policy or since}, meaning that either the policy holds or the alert is sounding.
The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}.
The \texttt{current} value is asserted when in the current time step there is a request with a response, or there is no request and no response.

The value of \texttt{previous} in the current time step relies on values from the previous time step.
The \texttt{->} operator designates initialization, as the previous time step is undefined in the first step of the system.
The left operand to the operator is the initial value of \texttt{previous} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system.
The right operand is the value of \texttt{previous} after the initial step.
Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step, \texttt{previous} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.

The value of \texttt{since} in \emph{Alert if start is not bounded relative to a request} relies on its own value in the previous time step.
The intuitive reading of the expression is that the alert has been true since the time when it first sounded.
The first \texttt{alrt} sets \texttt{since} to true, and once the value of \texttt{since} is true, that value persists as long as \texttt{alrt} holds.
The \emph{Alert if start is not bounded relative to a request} guarantee defines one requirement of a cyber-hardened system implementation.
Together with the other guarantees, the contract models the expected input and output of the system as a whole.

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.3]{dialogue.png}
    \end{tabular}
  \end{center}
  \caption{Wizard for automatically transforming the model with a filter.}
  \label{fig:dialogue}
\end{figure}

\newsavebox{\flt}
\begin{lrbox}{\flt}
\begin{lstlisting}[style=agree,numbers=left]
eq policy : bool = 
  WELL_FORMED_AUTOMATION_RESPONSE(Input);

guarantee Filter_Output
  "Filter output is well-formed" :
  if event(Input) and policy then 
    event(Output) and Output = Input
  else not event(Output);
\end{lstlisting}
\end{lrbox}

\newsavebox{\mntr}
\begin{lrbox}{\mntr}
\begin{lstlisting}[style=agree,numbers=left]
const is_latched : bool = true;
const MAX_LATENCY : int = 1;

eq rsp : bool = event(Response);
eq req : bool = event(Request);

eq isPending : bool = Since(not rsp, req and not rsp);
eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);

eq policy : bool = (rsp => req) ->
                   (    (isPending => latency < MAX_LATENCY)   
                    and (rsp => (req or pre(isPending))));
eq alert : bool = (not policy) -> 
                  ((is_latched and pre(alert)) or not policy);

assume "One outstanding request at a time" :
  (true -> (req => not pre(isPending)));

guarantee "Alert port tracks alert variable" :
  event(Alert) = alert;
guarantee "Output if not alerted" :
  if (not(alert) and rsp) then
    event(Output) and (Output = Response)
  else
    not (event(Output));    
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.62}{\usebox{\flt}} \\
    \end{tabular}
  \end{center}
  \caption{Contract specification for high-assurance filter.}
  \label{fig:filter}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.62}{\usebox{\mntr}} \\
    \end{tabular}
  \end{center}
  \caption{Contract specification for high-assurance monitor.}
  \label{fig:monitor}
\end{figure}

As noted previously, the original system fails to guarantee the cyber requirements.
BriefCASE provides two transformations to address the failing requirements: inserting a filter and inserting a monitor.
The component is added by selecting the connection in the model where the high-assurance component is to be added, and then choosing the appropriate transformation.
The system designer can provide transform configuration parameters in a wizard, as shown in \figref{fig:dialogue}.
The policy of the high-assurance component can be stated directly in the wizard, or it can be left blank.
In this example, the policy is specified as \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)}.
Additionally, because a transformation is ultimately driven by a cyber requirement, BriefCASE updates an embedded Resolute assurance case~\cite{resolute-destion}.
REMOVE: Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run at any time to determine whether those artifacts are valid.

The AGREE contract specification generated by the transform is shown in \figref{fig:filter}.
The guarantee is stylized for synthesis and completely defines the meaning of the output under every possible input.
The resulting AGREE specification for the monitor in this example is shown in \figref{fig:monitor}.
The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.
This behavior is one of the several options available in the dialogue.
The definition for \texttt{policy} is taken by the system developer from the contract in \figref{fig:sw}.
As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.