### 2022-03-19 paper/agree.tex
revised content added to semantics section. Checkpointing

This section gives an overview of AGREE's assume-guarantee reasoning
system, abstractly defining the verification conditions that AGREE
checks in order to establish overall system correctness at the model
level.  The discussion then becomes more concrete, showing how the
AGREE contract specification language is used to specify the
cyber-hardened system of \figref{fig:hardened}. It explains the the
top-level specification of the example system, then shifts to the
specifications for the filter and the monitor.  These two
specifications are those that are synthesized to equivalent CakeML by
SPLAT.

%% These conditions are useful in understanding how the semantics in
%% synthesis differ from those in the assume-guarantee reasoning
%% in \secref{sec:semantics}.

\subsection{Verification Conditions}
\input{verification-conditions}

\subsection{System Contract}
\input{sw-contract}

\subsection{Cyber-hardening the System Implementation}
\input{high-assurance-contracts}