### 2022-03-20 paper/agree.tex
added definition of code contract; other re-orgs and tweakage

This section gives an abstract overview of AGREE's compositional
reasoning system, defining the verification conditions that AGREE
checks in order to modularly establish overall system correctness at
the model level.  The discussion then becomes more concrete,
showing---by example---how the AGREE contract language is used to
specify the cyber-hardened system of \figref{fig:hardened}.  After the
system-level contract, discussion shifts to the filter and monitor
component specifications, which are structured to support code
generation.

%% These conditions are useful in understanding how the semantics in
%% synthesis differ from those in the assume-guarantee reasoning
%% in \secref{sec:semantics}.

\subsection{Model of computation}

The AGREE specification language is based on stream concepts, and
operators, from the Lustre language \cite{10.1145/41625.41641}.  Thus
the setting is synchronous dataflow where the inputs, outputs, and
expressions are data streams that comply with the input assumptions.
Contracts are evaluated in dependency order with inputs being
propagated to outputs through all contracts until they stabilize; as
such, the contracts, and thereby the top-level model, must be
acyclic. (An apparent syntactic cycle, where a component is linked
back to itself, may be broken temporally by inserting delay
elements.) \footnote{KLS: some undefined terms here: evaluation of
contracts, stabilization, synchronous step.}

Once the contracts have stabilized, the model takes a synchronous step
to the next input data in the input stream(s).  The semantics do not
model computation or communication delay.\footnote{KLS: how does a
delay element (mentioned above) work then?}  The output of one
contract is seen at the input of any downstream contract in the same
step of the input data stream.
\footnote{KLS: This subsection should give a helpful abstract description of what AG specifications are about, \ie, constraints on IO stream. Needs a rewrite, however.}

\subsection{Verification conditions}
\input{verification-conditions}

\subsection{System contract}
\input{sw-contract}

\subsection{Cyber-hardening the system implementation}
\input{high-assurance-contracts}