### 2022-03-23 paper/agree.tex
checkpointing. Re-org is done (for now). Mostly done describing AGREE contracts. Synthesis.tex merged with semantics.tex

This section gives an abstract mathematical overview of AGREE's
compositional reasoning system, defining assume-guarantee contracts
and the verification conditions that AGREE checks in order to
modularly establish overall system correctness at the model level.
The syntax and semantics of AGREE are sketched in some detail.  The
discussion then becomes more concrete, showing---by example---how the
AGREE contract language is used to specify the cyber-hardened system
of \figref{fig:hardened}. The discussion proceeds from the
system-level contract to the filter and monitor components, which rely
on the notion of a \emph{code contract} to support code generation.

\subsection{Contracts and verification conditions}
\input{verification-conditions}