### 2022-03-28 paper/sw-contract.tex

\newsavebox{\sw}
\begin{lrbox}{\sw}
\begin{lstlisting}[style=agree,numbers=left]
eq req : bool = event(AutomationRequest);*\label{line:sw-event-def-start}*
eq avl : bool = event(AirVehicleLocation);
eq wp  : bool = event(Waypoint);
eq strt: bool = event(Start);
eq alrt: bool = event(Alert);*\label{line:sw-event-def-end}*

assume "Automation requests are well-formed" : *\label{line:sw-assume-1}*
  req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
assume "Air vehicle locations are well-formed" : *\label{line:sw-assume-2}*
  avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);
assume "One automation request in flight at a time" : *\label{line:sw-assume-3}*
  true ->
  (req => pre(Historically(not req) or Since(not req, strt)));

guarantee "Waypoints coincide with air vehicle locations": *\label{line:sw-guarantee-1}*
  wp => avl;
guarantee "Starts include a new waypoint" : *\label{line:sw-guarantee-2}*
  strt => wp;
guarantee "Waypoints are well-formed" :  *\label{line:sw-guarantee-3}*
  wp => WELL_FORMED_WAYPOINT(Waypoint);
guarantee "Starts within one cycle of requests if not alerting" : *\label{line:sw-guarantee-4}*
  (strt => ((not alrt) and req)) ->
  (strt => ((not alrt) and (req or pre(req))));
guarantee "Alert if not started within one cycle of requests" : *\label{line:sw-guarantee-5}*
  true -> ((pre(req and not strt) and not strt) => alrt);
guarantee "Once alerted always alerted" : *\label{line:sw-guarantee-6}*
  Once(alrt) => alrt;
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \scalebox{0.62}{\usebox{\sw}}
  \end{center}
  \caption{The SW component contract.}
  \label{fig:sw}
\end{figure}

%% The actual surface syntax for AGREE that is embedded in OSATE is
%% slightly different than the formal semantics in the previous section.
%% The mapping between the two is straightforward.

%% The
%% specification uses \texttt{eq} statements to define variables local to
%% the contract specification.  For
%% example, \lineref{line:sw-event-def-start} defines the \texttt{req}
%% variable to be equivalent to the \texttt{event} expression.

%% All the named
%% ports in the corresponding AADL component are in the scope of the
%% specification, and there are additional implicit boolean \emph{event}
%% inputs (or outputs) associated with event ports.  An \texttt{event}
%% expression refers to that implicit input (or output) boolean value and
%% is true when data is present on the named port and false otherwise.

%% An \texttt{assume} statement is a string description followed by a
%% predicate.

%% When writing assumptions and guarantees,
%% it is important to differentiate pre-state, before a contract updates
%% its state, and post-state, after a contract updates its state, in
%% response to the current input.  The \texttt{pre} naturally makes that
%% distinction.  In general, assumptions regarding state should be
%% evaluated in the pre-state of the component with the current inputs,
%% and guarantees regarding state should be evaluated on the post-state
%% of the component given the current inputs.  Additionally, guarantees
%% should use \texttt{pre} anytime they need to reason about the current
%% state relative to the previous state.

%% This condition constrains when a request
%% can arrive by reasoning about the \emph{state} of the contract defined
%% by the associated streams. The assumption uses a followed-by
%% expression so it is \texttt{true} at time zero and then it is the
%% truth value of the implication in all future instances.  The
%% followed-by guards the \texttt{pre} as mentioned previously; thus
%% after time zero, the assumption depends on the presence of a request
%% and the pre-state of the component.  The assumption
%% on \lineref{line:sw-assume-3} is that

The AGREE specification for the SW component in
Section~\ref{sec:example} is given in \figref{fig:sw}.
\linesref{line:sw-event-def-start}{line:sw-event-def-end} define
local variables that are true when data is present on the
corresponding event ports for the component.  These variables are
purely for notational convenience. The contract has three assumptions:
\begin{itemize}
\item \lineref{line:sw-assume-1} and \lineref{line:sw-assume-2}.
When input is present it is well-formed. The well-formed predicates
are defined elsewhere using AGREE functions.

\item \lineref{line:sw-assume-3}. If there is an incoming request,
it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)},
or it arrives after the component has output a start event in response to a previous
request, \texttt{\textbf{Since}(\textbf{not} req, strt)}---to be read as \emph{not
request since start}.
\end{itemize}

The contract has six guarantees:
\begin{itemize}
\item \linesref{line:sw-guarantee-1}{line:sw-guarantee-3}
coincide events and assert well-formed output.\footnote{\konst{KLS}:I am confused about the ``coincide events''.}
\item \lineref{line:sw-guarantee-4} insists that a start happens with a request or one step after a request. At time zero the start must coincide
with the request; after that, the start must coincide with the request
or else is a response to a request one step earlier.  The guarantee
works with the assumption on \lineref{line:sw-assume-3} and thus does
not have to check for two requests in a row without a start, since the
assumption precludes that input behavior. The guarantee also does not
force the start to always happen, it only says that if it does happen,
it is under the defined conditions.

\item \lineref{line:sw-guarantee-5} forces the alert to sound if the start does not arrive within the one-step bound. Together with \lineref{line:sw-guarantee-4} this allows
for non-alerting and alerting behavior.
\item \lineref{line:sw-guarantee-6} guarantees that, if an alert has ever happened,
it continues to happen.
\end{itemize}
These six guarantees define the behavior of the SW component under the
three assumptions and correspond to the informal descriptions given
in \secref{sec:example}.