### 2022-03-29 paper/synthesis.tex
Done Section 7

Synthesis maps from model and specifications to code. The SPLAT tool
traverses the system architecture looking for occurrences of high
assurance components specified by code contracts; for each such
occurrence it generates a CakeML program. SPLAT supports three kinds
of behavioral specification from which to generate code. In increasing
order of expressiveness, these are: regular expressions, contiguity
types, and AGREE specifications. In earlier work we have described how
regular expressions and contiguity types are used as specification
languages for verified filter implementations. In the present paper,
we extend this \emph{specification oriented} approach to the
Turing-complete specifications of AGREE.  We give a short review of
that previous work before going into detail on AGREE.

\paragraph{Regular expressions.} Often the well-formedness of data on connections
can be enforced by matching against a regular expression. In other
words, the data can be characterized by a \emph{regular language}. As
is well known, regular expressions can be translated to efficient
deterministic finite state machine (DFA)
implementations. In \cite{formal-filter-synth-langsec,case-verified-filter}
we report on the creation and application of table-based DFA filters
from a verified regular expression-to-DFA compiler. The latter paper
offers a fully-worked out example synthesizing a regular expression
filter and formally proving that it meets an infinitary behavioral
specification.

\paragraph{Contiguity types\cite{contiguity-types}.}
Contiguity types, similar to regular expressions, are effective at
describing data at the network message level, \eg, as flat
strings. The notation is however much more expressive, being able to
declare many kinds of so-called \emph{self-describing} messages, \ie,
those where message structure is dependent on data held elsewhere in
the message. Figure \ref{fig:filter-spec} gives a contiguity type
specification for the welll-formed waypoints of our example system. We
have formalized and verified a contiguity type matcher, which has been
extensively applied throughout our case studies.

\subsection{Compiling from code contracts}

When full computational power is needed, we can use the expressive
power of AGREE code contracts to write arbitrarily complex filter or
monitor operations, as we have already seen. Our formal model of AGREE
syntax and semantics is applied to extract code contracts and map them
to equivalent logic functions. In particular, given a code contract with
functionality ${\Eqs}\cdot G$, input ports $\mathbb{I}$, state
variables $\mathbb{S}$, and output ports $\mathbb{O}$, the semantics
given in Section \ref{sec:code-contracts} can be applied to create a
function
\[
 \konst{strmStep} : \mathbb{N}_0 \to \mathit{Env} \to \mathit{Env} = \sem{{\Eqs}\cdot G}
\]
Assume ${\Eqs}\cdot G$ is \emph{Pascalish}, \ie, sequentially
imperative. Then we can direcctly create a corresponding logic function
\[
\konst{stateStep} : \mathit{input} \times \mathit{state} \to \mathit{state} \times \mathit{output}
\]
that takes input and state as parameters and returns an updated state
and output. Thus the environment $\mathit{Env}$ holding all previous
values of each variable can be replaced by \emph{parameters}, each of
which is able to hold only one data element. Assuming
$\mathit{Env}' = \konst{strmStep}\;t\; \mathit{Env}$, we can formally prove that
\[
\vdash (\mathit{Env}' \vert_{\mathbb{S}},
 \mathit{Env}' \vert_{\mathbb{O}}) =
\konst{stateStep} (\mathit{Env}\vert_{\mathbb{I}(t)},
                   \mathit{Env}\vert_{\mathbb{S}(t)})
\]
In other words, projecting out the state and output components from
$\konst{strmStep}\;t\; \mathit{Env}$ equals the result of
applying \konst{stateStep} on the input and state components of
$\mathit{Env}$ at time $t$. This theorem connects the stream semantics
of the code contract to a pure logic function over the values
appearing in the stream. From this point we can use tools in the
CakeML eco-system to jump from the setting of higher order logic to
CakeML abstract syntax \cite{cakeml-translator}. Thus we have moved
from stream semantics to logic function to CakeML evaluation semantics
in a property-preserving way. Once in CakeML, further necessary
features can be added, namely imperative variables for storing state
between component invocations, also invocations of the (verified) foreign
function interface of CakeML \cite{cakeml-monadic}. Finally, the CakeML compiler can be
applied to the program. It delivers an executable which can be shown
to preserve the original contract.

\subsection{Generating CakeML}
For any of the three specification languages, we take the following general steps to
arrive at an executable.

\begin{enumerate}
\item Map from specification to computable function in logic, producing
a proof of equivalence.
\item Add parsers and printers (specified by contiguity types) for input and output.
\item Map to CakeML and add calls to input/output library routines via
      the CakeML foreign function interface.
\item Invoke the CakeML compiler.
\end{enumerate}

%% Starting from the semantics of the code contract described in
%% Section \ref{sec:code-contracts} a sequence of semantics-preserving
%% steps are taken:
%% \begin{itemize}
%% \item $\sem{{\Eqs}\cdot G}$ yields a function on arbitrary depth streams
%% \item transform to Pascalish
%% \item transform to logic function
%% \item attach buffer handling to inputs and outputs (contig type parsing)
%% \item translate to CakeML
%% \item run CakeML compiler
%% \end{itemize}

In the following, we examine further details in both filter and
monitor synthesis.

\subsection{Filter Generation}

A filter is intended to be simple, although it may make deep semantic
checks. A filter has one input port and one output; messages on the
input that the filter policy admits pass unchanged to the output port;
all others are dropped (not passed on). We have investigated two kinds
of filter. In the first, a relatively shallow scan of the input
suffices to enforce the policy. For example, we have used the
expressive power of regular expressions and Contiguity
Types \cite{contiguity-types} to enforce \emph{lightweight} bounds
constraints on GPS coordinates in UxAS messages. On the other hand, a
filter may need to parse the input buffer into a data structure
specified in AGREE and apply a user-defined \emph{wellformedness}
property, also specified in AGREE, to the data. Arbitrarily complex
wellformedness checks can be made in this
way. \figref{fig:filter-spec} shows a combination where the checking
specified by {\small\verb+WELL_FORMED_AUTOMATION_RESPONSE+} depends on
an underlying check specified by the contiguity type checking bounds
on waypoints.

The verdict of a filter is made and performed within one thread
invocation. Thus, in its given time slice, the following steps must be
completed:

\begin{enumerate}

\item The filter checks to see if there is any input available.  If there is none
then it yields control; otherwise:

\item The input is read (and parsed if need be);

\item The wellformedness predicate is evaluated on the input;

\item If the predicate returns \konst{true} then the input buffer
 is copied to the output, otherwise no action is taken; and

\item The filter yields control.
\end{enumerate}

\newsavebox{\contig}
\begin{lrbox}{\contig}
\begin{lstlisting}[style=myML]
  Waypoint =
    {Latitude  : f64
     Longitude : f64
     Altitude  : f32
     Check     : Assert
      (~90.0 <= Latitude and Latitude <= 90.0 and
       ~180.0 <= Longitude and Longitude <= 180.0 and
       1000.0 <= Altitude and Altitude <= 15000.0)}

  AutomationResponse =
    {TaskID : i64
     Length : u8
     Waypoints : Waypoint [3]}

 fun WELL_FORMED_AUTOMATION_RESPONSE(aresp) =
   (forall wpt in aresp.Waypoints, WELL_FORMED_WAYPOINT(wpt))
   and ... ;
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\contig}}
    \end{tabular}
  \end{center}
  \caption{Filter specification.}
  \label{fig:filter-spec}
\end{figure}

\newsavebox{\cml}
\begin{lrbox}{\cml}
\begin{lstlisting}[style=myML]
fun filter_step () =
 let val () = Utils.clear_buf buffer
     val () = API.callFFI "get_input" "" buffer
 in
    if WELL_FORMED_AUTOMATION_RESPONSE buffer
    then
      API.callFFI "put_output" buffer Utils.emptybuf
    else print"Filter rejects message.\n"
end
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\cml}}
    \end{tabular}
  \end{center}
  \caption{Synthesized CakeML for the filter.}
  \label{fig:filter-cakeml}
\end{figure}

The contiguity type specification and wellformedness predicate for
the filter are shown in \figref{fig:filter-spec} and the synthesized
CakeML code is in \figref{fig:filter-cakeml}. The code is called at
dispatch by the scheduler. The \texttt{API.callFFI} is the link to the
communication fabric to capture input and provide output. The body of
the function restates the filter contract to make the appropriate
assignments in a way that matches the truth value of the predicate in
the filter guarantee.  The auto-generated AGREE specification raises
an alert output when the relation is violated.

\subsection{Monitor Generation}

%% Monitors are intended to track and analyze the externally visible
%% behavior of system components through time. Therefore, they tend to
%% require more extensive computational ability than filters. In
%% particular, our basic notion of a monitor is that it embodies a
%% relation over its input and output streams, and is able to access the
%% value of a stream at any earlier point in time, if necessary. Monitors
%% commonly use state to keep track of earlier values, unlike filters
%% which, for us, are typically stateless components. (However, there is
%% nothing in our approach that forbids stateful filters: they can be
%% realized by monitors.)
As discussed above, a monitor specification is
mapped to a state transformation function
%
\[
\konst{stateStep} : \mathit{input} \times \mathit{state} \to \mathit{state} \times \mathit{output}.
\]
which computes as follows:
\begin{enumerate}

\item Each input buffer is parsed into data of the type specified by the port
  type;

\item New values for the state variables are computed, in dependency
  order. At this point variables defined by \emph{followd-by} are
  dealt with. Suppose the variable definitions have the following
  form:
\[
\begin{array}{l}
  v_1 = i_1 \longrightarrow e_1 \\
  \cdots \\
  v_n = i_n \longrightarrow e_n \\
\end{array}
\]
In the generated function, for the first invocation of \konst{stateStep} only,
the initializations are executed in order:
\[
  v_1 = i_1; \cdots ;  v_n = i_n;
\]
In all later invocations, the \emph{non-initialization} assignments are performed, also in order:
\[
  v_1 = e_1; \cdots ; v_n = e_n; \\
\]

\item Values of the outputs are computed;

\item Outputs are written and the new state is written;

\item The monitor yields control.
\end{enumerate}

The \konst{stepFn} for the monitor of the example described in
Section~\ref{sec:example} is displayed in \figref{fig:monitor-cakeml}.

\newsavebox{\monFn}
\begin{lrbox}{\monFn}
\begin{lstlisting}[style=myML]
fun stateStep (response,request)
    (rsp,req,preIsPending,isPending,latency,policy,alert)
let val stateVars' =
      if !initStep then
      let val rsp = Option.isSome response
          val req = Option.isSome request
          val preIsPending = False
          val isPending = req andalso not rsp
          val latency = 0
          val policy = (if rsp then req else True)
          val alert = not(policy)
          val () = (initStep := False)
      in
        (rsp,req,preIsPending,isPending,latency,policy,alert)
      end
      else
      let ...
          val rsp = Option.isSome response
          val req = Option.isSome request
          val preIsPending = pre isPending
          val isPending = (req andalso not rsp) orelse
                          ((not rsp) andalso pre isPending)
          val latency = (if req then 0 else (pre latency + 1)
          val policy =
            ((if isPending then
                (latency < Defs.max_latency) else True)
              andalso
              (if rsp then
                 (req orelse preIsPending) else True))
          val alert = (Defs.is_latched andalso (pre alert))
                      orelse not(policy)
      in
        (rsp,req,preIsPending,isPending,latency,policy,alert)
      end
   val (rsp',_,_,_,_,_,alert') = stateVars'
   val alertPort = if alert' then Some () else None
   val outputPort =
     if alert' then None else
     if rsp' then Some response
     else None
in
   (stateVars', (alertPort,outputPort))
end
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\monFn}}
    \end{tabular}
  \end{center}
  \caption{Synthesized CakeML for the monitor.}
  \label{fig:monitor-cakeml}
\end{figure}

\subsection{Incorporation into system execution}

The system scheduler \emph{activates} components in some order. It is
an obligation on the system that the scheduler follows some sensible
partial order of component activation and allows each component
sufficient time for its computation.  Activating a monitor component
takes the form of the following pseudo-code, in which the monitor
evaluates the \konst{stateStep} on its current inputs and the current
values of the state variables, returning the new state and the output
values.
\[
\begin{array}{ll}
 \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
 (v_1,\ldots) & = \konst{readState}() ; \\
 ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stateStep} ((i_1,\ldots),(v_1,\ldots)) ; \\
 \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
 \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
\end{array}
\]

\paragraph{Partiality.} Partiality is an important consideration for both filters and
monitors: the input data might not be parseable or the wellformedness
computation of a filter could be badly written and fail at runtime. In
such cases, the filter should recover and yield control without
passing the input onwards. For these failure cases, the filter is
behaving as it should, but we must also guard against situations in
which a \emph{correctly specified} filter or monitor fails at
runtime. This kind of defect arises when a filter, for
example, \emph{ought} to accept a message, but lack of resources
results in the filter failing to do so. For example, the parse of a
message might need more space than has been allocated; another example
could be if the time slice provided by the scheduler is too short for
the wellformedness computation to finish. Thus resource bounds need to
be included in the correctness argument. Some preliminary work that
could support formal proofs of such bounds has been done for
CakeML \cite{cakeml-space-cost}.

%% \subsection{Component Behavior}

%% Intuitively, for monitor specification $s$, \konst{stepFn} is the
%% concrete embodiment of $\konst{SynthEval}\;s$, as defined in Section
%% \ref{agree-semantics}. Its correctness amounts to showing that, given
%% a sequence of inputs, and an initial state meeting the initialization
%% constraints, iterating \konst{stepFn} produces a $\pi$ s.t. $\pi \in
%% \Lang{s}$; and taking the union over all input sequences and
%% initial states produces $\Lang{s}$ itself.

%% \subsubsection{Initialization}

%% A monitor may need to accumulate a certain minimum number of
%% observations before being able to make a meaningful assessment of
%% behavior. Until that threshold is attained, the monitor is essentially
%% in its \emph{initialization} phase. In order for correct code to
%% be generated, monitor specifications need to spell out the values of
%% output ports when in their initialization phases. For example, suppose a
%% monitor does some kind of differential assessment of inputs at
%% adjacent time slices, alerting when (say) the measured location of a
%% UAV at times $t$ and $t+1$ is such that the distance between the two
%% locations is unusually large. Such a monitor needs two measurements
%% before making its first judgement, but at the time of its first
%% output, only one measurement will have been made. The specification
%% must then explicitly state the correct value for the first output.

% \subsubsection{Step function}