### 2022-09-15 paper/intro.tex

In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  
Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.

Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
They instead rely on process-oriented checklists and guidelines.
Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation. This is not a sustainable development model.

The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).  The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S. Government.}
One such tool is the Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

This manuscript details how \brfcs\ facilitates inserting, specifying, testing, and synthesizing high assurance components into a system to improve its cyber-resiliency.
These are referred to as \emph{transforms} by \brfcs.
Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components.
\brfcs\ automates inserting the components into the system, and the designer writes a \emph{code contract} for each component to define its intended behavior.

A code contract is a formal specification in the Assume Guarantee Reasoning
Environment (\agr) language.
\agr is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
The code contract language is Turing complete allowing the designer to specify arbitrarily complex behavior.
These contracts are unit tested for correctness with \emph{test contracts}.
Test contracts define test scenarios to be implemented by the code contract under test. 
\agr\ proves whether or not that code contract implementation is correct enabling to designer to iteratively test the component behavior inside the \brfcs\ environment.
Once the behavior of the code contract is verified, \agr proves that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

Another novel aspect of the approach discussed in this manuscript is the synthesis of the code contracts for the high-assurance components to \ckml, a verified
compiler implementation for the functional programming language
ML \cite{cakeml}. This manuscript describes in detail the synthesis path
from code contracts to \ckml\ code, providing a formal framework in
which to argue correctness. \ckml\ then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the \ckml\ source code is exactly
preserved in the final binaries). 
A key contribution here is that the code contract semantics are defined in such a way that the \agr\ verification 
results for the code contract hold for the deployed component, \ie, the component will detect
and prevent the indicated cyber-vulnerabilities over all possible
finite inputs. Preliminary work has shown how to lift this result to
infinite input traces as these systems are inherently reactive and
intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

The manuscript further details a 
case study applying these transformations with \brfcs\ to an Unmanned Aerial
Vehicle (UAV) system that uses the Air Force Research Laboratory's
OpenUxAS services for route planning.
OpenUxAS, as an open source product, is considered \emph{untrusted}.
The UAV system is thus transformed to be resilient to malicious behavior that may arise in the untrusted component.
Here the transforms add filters to guard
against malformed input and monitors to guard against malicious flight plans from OpenUxAS. The case study
system is complex and shows
the viability of the approach in potential full-scale industrial
design.

The \brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
Our approach currently applies to common cyber-vulnerabilities, such
as overflow, lack of input validation, and supply-chain issues; however, other cyber-vulnerabilities such as
side-channel attacks and denial of service are not yet dealt with in
our work.  Here, we do not report on the invention of a new type of
high-assurance component in terms of capability; instead, our contribution
is in the automated synthesis of security-improving components from
formal specifications and a means to show that the synthesis is
correct.

The approach is illustrated by a simple example in
\secref{sec:example}. \agr\ verification and specification are
presented in \secref{sec:agree} and \secref{agree-semantics}.
Code contracts are defined in \secref{sec:code-contracts} with a method
for testing code contracts in the \agr\ framework before synthesis given
\secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}.

\begin{comment}
  BriefCASE incorporates model-level cyber analysis tools (presently
  GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
  examine AADL models for potential vulnerabilities and suggest
  cyber-security requirements to mitigate them.  A library of
  architectural transforms guides the system engineer through automated
  model transformations that modify the architecture to address these
  requirements, possibly inserting new high-assurance components into
  the system.  Implementations for the new components are synthesized
  from formal specifications using
  SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
  (Semantic Properties for Language and Automata Theory).

  Formal
  verification that the transformed system model meets its requirements
  is accomplished via \agr~\cite{agree2013} (Assume Guarantee Reasoning
  Environment).
  %The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
  \agr\ is a {\em compositional assume-guarantee} style model checker
  for AADL models that attempts to prove properties about one layer of
  an architecture using properties allocated to its subcomponents.
  Cyber-resilient code implementing the verified model is then
  automatically generated using the High Assurance Modeling and Rapid
  Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
  desired, this code can be targeted to the formally verified seL4
  secure microkernel~\cite{sel4-2009}.
\end{comment}