### 2022-09-17 cover-letter/resubmit-letter.tex
Add letter summarizing review responses

\documentclass[10pt]{byu-cs}

\begin{document}
\pagestyle{headings}

\signature{Eric G Mercer\\
           Associate Professor}

\begin{letter}{MoDELS 2021 Special Issue Resubmission}

\date{\today}

\opening{Shiva Nejati and D\'{a}niel Varr\'{o},}

On behalf of my co-authors, I want to thank the reviewers for the very thoughtful feedback on the first version of our paper.
We have carefully studied that feedback, made appropriate revisions, and are hopeful to have responded to the reviewer comments in a satisfactory way.
We feel that those comments have resulted in a much improved submission.

To aid in reviewing this new manuscript, I am including here a detailed summary of the changes in this new submission.
\begin{itemize}
  \item We revised the introduction to make clear the problem being solved relative to existing work and state the novelty of our contribution. It has also been expanded to address upfront the other requests, highlighted by the editor, in the prior review to (1) describe the dependency on the chosen language and tools; (2) define the limitations of the approach, and (3) discuss how this approach could be extended to other domains.
  \item We added a new section summarizing BriefCASE, with its many tools, and its general workflow. That section additionally makes clear the parts of BriefCASE that relate to our contributions in the paper.
  \item We added a new workflow diagram to the section discussing the illustrative example to show the inputs, outputs, existing adopted tools, and new tools that are part of the contribution of this work. We also removed from the example anything in BriefCASE that is not directly related to the contributions discussed in this submission.
  \item We have grouped into a section titled "Background" everything relating to the AGREE tool as it is an existing tool that we leverage as part of our solution to the synthesis problem.
  \item We have revised the case study to identify roles in identifying and adding requirements.
  \item We have completely revised the related work to contrast our solution to synthesis relative to the existing solutions with an effort to highlight that we have solved the problems identified in the introduction. We have also added new related work as suggested by the reviewers and other related work that we have became aware of since submitting our original paper.
  \item We have revised the conclusion to be specific in the planned and ongoing future work as requested by the reviewers. In particular, we have detailed how we plan to fully verify the transformation from a code contract to CakeML and reported on some early results in that regard. We have also referenced ongoing useability studies with Collins designers.
  \item We have made several edits through the document for consistency and to address other specific reviewer comments.
\end{itemize}

I hope that in reviewing this new submission that satisfied with how we have addressed the feedback in the prior reviews.
We thank you for your consideration and the valuable feedback.

\closing{Sincerely,}

\end{letter}
\end{document}