### 2022-09-17 paper/intro.tex

\egm{Add to introduction and related work, that what is done at
Galois called Copilot. Its a stream processing langague that generates
code from the specification. Copilot to C. Effectively CodeGen from
Lustre. Bounds how far needed to look in past for any value.}

In recent years, aerospace stakeholders have realized that avionics
systems are subject to possible cyber-attacks just like other
cyber-physical systems.
%
\footnote{This work was funded in part by the
Defense Advanced Research Projects Agency (DARPA).  The views
expressed are those of the authors and do not reflect the official
policy or position of DARPA or the U.S. Government.}
%
Thus, in addition to being fault-tolerant, safety-critical avionics
systems must also be {\em cyber-resilient}.  Cyber-resiliency means
that the system is tolerant to cyber-attacks just as safety-critical
systems are tolerant to random faults: they recover and continue to
execute their mission function, or safely shut down, as requirements
dictate.

Unfortunately, systems engineers are currently given few development
tools to help answer even basic questions about potential
vulnerabilities and ways to mitigate vulnerabilities.  They instead
rely on process-oriented checklists and guidelines.  Cyber
vulnerabilities are often discovered during penetration testing late
in the development process; or worse yet, they may be discovered only
after the product has been fielded, necessitating extremely expensive
and time-consuming remediation. This is not a sustainable development
model.

%% The DARPA Cyber Assured Systems Engineering (CASE) project is targeted
%% at developing tools for design, analysis, and verification that enable
%% systems engineers to {\em design-in} cyber-resiliency for complex
%% cyber-physical systems.

We have been developing the {\em BriefCASE} toolsuite to address this need.
\brfcs\ is a Model-Based Systems Engineering (MBSE) environment
built in the Open Source AADL\footnote{AADL is the acronym for
Architecture Analysis and Design Language~\cite{aadl}.}  Tool
Environment (OSATE) to add new design, analysis, and code generation
capabilities for building cyber-resilient systems.

In this paper we describe how \brfcs\ facilitates inserting,
specifying, testing, and synthesizing high assurance components into a
system to improve its cyber-resiliency.  The main organizing concept
is that of an architecture-to-architecture \emph{security-improving
transform}, achieved via the insertion of a new architectural
component aimed at mitigating a cyber-vulnerability.  We describe two
cyber-resiliency transformations in this paper: (1) the insertion of a
filter to prevent malformed data from a malicious actor being
propagated to downstream components, and (2) the insertion of a
runtime monitor to detect (and alert) unexpected behaviors arising
from untrusted components.

A code contract is a formal specification in the Assume Guarantee
Reasoning Environment (\agr) language.
\agr\ is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
The code contract language is Turing complete allowing the designer to
specify arbitrarily complex behavior.  These contracts are unit tested
for correctness with \emph{test contracts}.  Test contracts define
test scenarios to be implemented by the code contract under test.
\agr\ proves whether or not that code contract implementation is correct, enabling the designer to iteratively test the component behavior inside the \brfcs\ environment.
Once the behavior of the code contract is verified, \agr\ proves that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

Another novel aspect of our workthe approach discussed in this manuscript is
the synthesis of the code contracts for the high-assurance components
to \ckml, a verified compiler implementation for the functional
programming language ML \cite{cakeml}.  This manuscript describes in
detail the synthesis path from code contracts to \ckml\ code,
providing a formal framework in which to argue correctness. \ckml\
then provides a verified compilation path to several different target
binaries (and also proving that the meaning of the \ckml\ source code
is exactly preserved in the final binaries).  A key contribution here
is that the code contract semantics are defined in such a way that
the \agr\ verification results for the code contract hold for the
deployed component, \ie, the component will detect and prevent the
indicated cyber-vulnerabilities over all possible finite
inputs. Preliminary work has shown how to lift this result to infinite
input traces as these systems are inherently reactive and intended to
run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

The manuscript further details a case study applying these
transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV)
system that uses the Air Force Research Laboratory's OpenUxAS services
for route planning.  OpenUxAS, as an open source product, is
considered \emph{untrusted}.  The UAV system is thus transformed to be
resilient to malicious behavior that may arise in the untrusted
component.  Here the transforms add filters to guard against malformed
input and monitors to guard against malicious flight plans from
OpenUxAS. The case study system is complex and shows the viability of
the approach in potential full-scale industrial design.

\brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
Our approach currently applies to common cyber-vulnerabilities, such
as overflow, lack of input validation, and supply-chain issues;
however, other cyber-vulnerabilities such as side-channel attacks and
denial of service are not yet dealt with in our work.  Here, we do not
report on the invention of a new type of high-assurance component in
terms of capability; instead, our contribution is in the automated
synthesis of security-improving components from formal specifications
and a means to show that the synthesis is correct.

We now give a summary of the contributions detailed in this
manuscript.
\begin{compactitem}
  \item The language and semantics of code contracts to specify the behavior of high assurance components.
  \item Test contracts as a stylized mechanism to unit test code contracts inside the \brfcs\ framework with \agr.
  \item A synthesis path from code contracts to \ckml\ with a formal framework to argue correctness.
  \item A formal argument that \agr\ verification of code contracts carry over to the resulting binaries from the \ckml\ compiler.
  \item A case study that used the implementation of the approach in \brfcs\ to add cyber-resiliency to a non-trivial UAV system.
\end{compactitem}
The rest of this manuscript is organized as follows.
\secref{sec:overview} is an overview of the \brfcs\ environment and key tools in that environment relative to the contributions of this work. The approach is illustrated by a simple example in
\secref{sec:example}. Essential background on \agr\ verification and specification are
presented in \secref{sec:agree}.
The language and semantics of code contracts are defined in \secref{sec:code-contracts}, and test contracts are defined in \secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}.
\secref{sec:case-study} discusses the case study.
This is followed by related work in \secref{sec:related-work}.
The conclusions and future work are in \secref{sec:conclusion}.

\begin{comment}
  BriefCASE incorporates model-level cyber analysis tools (presently
  GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
  examine AADL models for potential vulnerabilities and suggest
  cyber-security requirements to mitigate them.  A library of
  architectural transforms guides the system engineer through automated
  model transformations that modify the architecture to address these
  requirements, possibly inserting new high-assurance components into
  the system.  Implementations for the new components are synthesized
  from formal specifications using
  SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
  (Semantic Properties for Language and Automata Theory).

  Formal
  verification that the transformed system model meets its requirements
  is accomplished via \agr~\cite{agree2013} (Assume Guarantee Reasoning
  Environment).
  %The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
  \agr\ is a {\em compositional assume-guarantee} style model checker
  for AADL models that attempts to prove properties about one layer of
  an architecture using properties allocated to its subcomponents.
  Cyber-resilient code implementing the verified model is then
  automatically generated using the High Assurance Modeling and Rapid
  Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
  desired, this code can be targeted to the formally verified seL4
  secure microkernel~\cite{sel4-2009}.
\end{comment}