### 2022-12-03 cover-letter/refactor-letter.tex
Revise abstract as per reviewer feedback

\documentclass[10pt]{byu-cs}

\begin{document}
\pagestyle{headings}

\signature{Eric G Mercer\\
           Associate Professor}

\begin{letter}{MoDELS 2021 Special Issue Resubmission}

\date{\today}

\opening{Shiva Nejati and D\'{a}niel Varr\'{o},}

On behalf of my co-authors, I want to thank the reviewers for the very thoughtful feedback on the second version of our paper.
We have carefully studied that feedback, again made appropriate revisions, and are hopeful to have responded to the reviewers' feedback in a satisfactory way.
We feel that the feedback has improved the submission.

To aid in reviewing this new refactored manuscript, I am including here a detailed summary of the changes, and where possible, the submission itself indicates the changes directly to make it easier for the reviewers to inspect.
\begin{itemize}
  \item We rewrote the abstract to be active present tense, define the problem, approach, contributions, and results.
\end{itemize}

I hope this new submission satisfies the reviewers' requests.

We thank you for your consideration and the valuable feedback.

\closing{Sincerely,}

\end{letter}
\end{document}