### 2022-12-19 cover-letter/refactor-letter.tex
Add transition paragraph to the AGREE background.

\documentclass[10pt]{byu-cs}

\begin{document}
\pagestyle{headings}

\signature{Eric G Mercer\\
           Associate Professor}

\begin{letter}{MoDELS 2021 Special Issue Resubmission}

\date{\today}

\opening{Shiva Nejati and D\'{a}niel Varr\'{o},}

On behalf of my co-authors, I want to thank the reviewers for the very thoughtful feedback on the second version of our paper.
We have carefully studied that feedback, again made appropriate revisions, and are hopeful to have responded to the reviewers' feedback in a satisfactory way.
We feel that the feedback has improved the submission.

To aid in reviewing this new refactored manuscript, I am including here a detailed summary of the changes, and where possible, the submission itself indicates the changes directly to make it easier for the reviewers to inspect.
\begin{itemize}
  \item We rewrote the abstract to be active present tense, define the problem, approach, contributions, and results.
  \item We revised the introduction to pull in the last three paragraphs of section two, and to better funnel readers from the big picture, intuitive overview, to specific details on our tools and contributions. 
  \item We added a discussion section to the case study in section 8 to address lessons learned from our experience. We also enlarged significantly figures 14 and 17.
  \item We added a paragraph to the end of the example section explaining that our discussion of AGREE early in the paper, of necessity, is a birdseye view of its role in the BriefCASE workflow for the reader that perhaps doesn't want all the detail in the following sections. The actual AGREE section of the paper then presents notation, semantics, and usage so that the reader is able to contextualize our primary contributions. We also added in that paragraph direct references to the sections in the paper where the detailed contracts for the example system are discussed. 
\end{itemize}

I hope this new submission satisfies the reviewers' requests.

We thank you for your consideration and the valuable feedback.

\closing{Sincerely,}

\end{letter}
\end{document}