### 2022-12-20 cover-letter/refactor-letter.tex
Revise the cover-letter

\documentclass[10pt]{byu-cs}

\begin{document}
\pagestyle{headings}

\signature{Eric G Mercer\\
           Associate Professor}

\begin{letter}{MoDELS 2021 Special Issue Resubmission}

\date{\today}

\opening{Shiva Nejati and D\'{a}niel Varr\'{o},}

On behalf of my co-authors, I want to thank the reviewers for the very thoughtful feedback on the second version of our paper.
We have carefully studied that feedback, again made appropriate revisions, and are hopeful to have responded to the reviewers' feedback in a satisfactory way.
We feel that the feedback has improved the submission.

To aid in reviewing this new refactored manuscript, I am including here a detailed summary of the changes, and where possible, the submission itself indicates the changes directly to make it easier for the reviewers to inspect by striking out old text and indicating new text in red. These annotations are easily removed for the camera version should no further edits be required by the reviewers. 
\begin{itemize}
  \item We rewrote the abstract to be active present tense, define the problem, approach, contributions, and results.
  \item We revised the introduction to pull in the last three paragraphs of section two, and to better funnel readers from the big picture, intuitive overview, to specific details on our tools and contributions. 
  \item We added a discussion section to the case study in section 8 (now section 7) to address lessons learned from our experience. We also enlarged significantly figures 14 and 17.
  \item We removed the overview section, created a new figure showing the workflow through the BriefCASE tool, and integrated that into the example section. We added to the example section references to the workflow figure. We moved to the beginning of the section a better overview of the example UAV system, and we added text explaining why the system must be cyber-resilient. We did not remove sections 3.1 - 3.3 (now sections 2.2 to 2.4) as suggested by one of our reviewers as these sections discuss our contributions at a high-level in the context of our illustrative example. We did verify that we had appropriate references to the example, and its various figures, in those sections as suggested by the same reviewer.
  \item We verified that sections 4 through 7 (now sections 3 through 6) have appropriate references to the example system.
  \item We added a paragraph to the end of the example section explaining that our discussion of AGREE early in the paper, of necessity, is a birdseye view of its role in the BriefCASE workflow. The actual AGREE section of the paper then presents notation, semantics, and usage so that the reader is able to contextualize our primary contributions. We also added in that paragraph, as well as in the presentation of the example in section 2, direct references to the sections, and figures, in the paper where the detailed contracts for the example system are discussed. 
  \item We made various other stylistic changes to improve the flow of the paper.
\end{itemize}

I hope this new submission satisfies the reviewers' requests.

We thank you for your consideration and the valuable feedback.

\closing{Sincerely,}

\end{letter}
\end{document}